Bayesian Sparsification of $\cplx$-valued networks

### 2019-11-12 paper/paper.tex
fixed formula overflows

\textit{Stochastic Gradient Variational Bayes} (SGVB), proposed by \citep{kingma_auto-encoding_2014},
is a differentiable unbiased Monte-Carlo estimator of \eqref{eq:elbo_general}, which makes
it possible to employ stochastic gradient-based optimization methods. If $
  \omega \sim q_{\theta}
$ is equivalent in distribution to $
  \omega = g(\varepsilon; \theta)
$ for some non-parametric random variable $
    \varepsilon \sim p_\varepsilon
$ and differentiable $
  g(\varepsilon; \theta)
$, then the SGVB is
\begin{equation}  \label{eq:sgvb}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta} \| \pi_{\lambda})
      % + \frac{N}{\lvert B \rvert} \sum_{i\in B}
        % \log p(z_i \mid \omega_i)b
      + \frac1{L} \sum_{l=1}^L
        \log p_{\phi}(D \mid g(\varepsilon_{l}; \theta))
        % \log p_{\phi}(D \mid \omega_{l})
        %   \Big\vert_{\omega_{l} = g(\varepsilon_{l}; \theta)}
      % =
      % - KL(q_{\theta} \| \pi)
      % + N \mathbb{E}_{B \sim \{D\}_M \otimes q_\theta^M}
        % \hat{\mathbb{E}}_{z,\omega \sim B}
          % \log p(z \mid \omega)
    \,, 
\end{equation}
where $
  (\varepsilon_{l})_{l=1}^L
$ is an iid sample from $p_\varepsilon$. Intractable KL-divergence terms can be replaced
by differentiable unbiased Monte-Carlo estimators, \citep{kingma_auto-encoding_2014}.
%
For the purposes of this study, we assume that the observed data is independent, and
use minibatch version of \eqref{eq:sgvb} with $L=1$:
\begin{equation}  \label{eq:elbo}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta} \| \pi_{\lambda})
      + \frac{N}{M} \sum_{i=1}^M
        \log p_{\phi}(z_i \mid g(\varepsilon_i; \theta))
    \,,
\end{equation}
for a random subsample $(z_i)_{i=1}^M$ from $D$ and iid $
  (\varepsilon_i)_{i=1}^M \sim p_\varepsilon
$.
%
\todo{DSVI}

@@

In particular, linear layers
$
  L \colon \cplx^{\mathrm{[in]}}
    \to \cplx^{\mathrm{[out]}}
$ act upon their inputs as follows:
\begin{equation}  \label{eq:cplx-lin-op}
  \real^{2 \times \mathrm{[in]}}
    \to \real^{2 \times \mathrm{[out]}}
    \colon (x_r, x_i)
      \mapsto \Bigl(
        L_r x_r - L_i x_i,
        L_r x_i + L_i x_r
      \Bigr)
    \,,
\end{equation}
where $
  L_r, L_i
    \colon \real^{\mathrm{[in]}}
      \to \real^{\mathrm{[out]}}
$ are unique operators that make up $L = L_r + j L_i$. A $\cplx$-convolutional layer with
kernel $W$ can be implemented as two $\real$-convolutions with kernels $\Re{W}$ and $\Im{W}$
using \eqref{eq:cplx-lin-op}. Algebraic operations, trigonometric functions, are implemented
through operations in $\real^2$. Activations, considered in \citep{trabelsi_deep_2017} include
independent componentwise application of $\real$-valued non-linearities, $
  z = u + j v
    \mapsto \sigma(u) + j \sigma(v)
$, $\cplx$-differentiable exponential and trigonometric activations, and \textit{modRelu}
activation, which applies soft-thresholding to the modulus and preserves the angle: $
  z = r e^{j \phi}
    \mapsto e^{j \phi} (r - b)_+
  % z \mapsto z \bigl(
  %     1 - \tfracb{\lvert z \rvert}
  %   \bigr)_+
$ with a learnable threshold parameter $b \in \real$.

@@

  \begin{pmatrix}

@@

  \end{pmatrix}

@@

      \end{pmatrix},
      \frac12
      \begin{pmatrix}
        \Re{(\Gamma + C)} & - \Im{(\Gamma - C)} \\

@@

From \eqref{eq:cn-paired-real-density} the entropy of a $\cplx$-Gaussian in terms of
$\Gamma$ and $C$ is

@@

A $\cplx$-Gaussian vector $z$ is \textit{proper}, if $z$ and $\conj{z}$ are uncorrelated, i.e.
$C = 0$. The entropy of proper $\cplx$-Gaussians is

@@

    \tag{\ref{eq:c-gauss-entropy-derivation}'}\,.

@@

In $\cplx$-variational dropout we use mean-field proper $\cplx$-Gaussian variational
approximation. Specifically, parameters $W \in \cplx^{n\times m}$ of a dense layer follow
\begin{equation}  \label{eq:c-gauss-vi}
  % (we omit the subscript $m=1$ for brevity)
  W \sim q(W)
    = \prod_{ij} q(W_{ij})
  \,,\,
    q(W_{ij})
      = \mathcal{N}^{\cplx}(W_{ij} \mid \mu_{ij}, \sigma^2_{ij}, 0)
  \,,
\end{equation}
and we use similar factorized approximation for convolution. We propose to MAP estimates,
of the biases i.e. assume an independent fully factorized Dirac approximation.

>>>>

\textit{Stochastic Gradient Variational Bayes} (SGVB), proposed by \citep{kingma_auto-encoding_2014},
is a differentiable unbiased Monte-Carlo estimator of \eqref{eq:elbo_general}, which makes
it possible to employ stochastic gradient-based optimization methods. If $
  \omega \sim q_{\theta}
$ is equivalent in distribution to $
  \omega = g(\varepsilon; \theta)
$ for some non-parametric random variable $
    \varepsilon \sim p_\varepsilon
$ and differentiable $
  g(\varepsilon; \theta)
$, then the SGVB is
\begin{equation}  \label{eq:sgvb}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta} \| \pi_{\lambda})
      % + \frac{N}{\lvert B \rvert} \sum_{i\in B}
        % \log p(z_i \mid \omega_i)b
      + \frac1{L} \sum_{l=1}^L
        \log p_{\phi}(D \mid g(\varepsilon_{l}; \theta))
        % \log p_{\phi}(D \mid \omega_{l})
        %   \Big\vert_{\omega_{l} = g(\varepsilon_{l}; \theta)}
      % =
      % - KL(q_{\theta} \| \pi)
      % + N \mathbb{E}_{B \sim \{D\}_M \otimes q_\theta^M}
        % \hat{\mathbb{E}}_{z,\omega \sim B}
          % \log p(z \mid \omega)
    \,, 
\end{equation}
where $
  (\varepsilon_{l})_{l=1}^L
$ is an iid sample from $p_\varepsilon$. Intractable KL-divergence terms can be replaced
by differentiable unbiased sampling estimators, \citep{kingma_auto-encoding_2014}.
%
For the purposes of this study, we assume that the observed data is independent, and
use minibatch version of \eqref{eq:sgvb} with $L=1$:
\begin{equation}  \label{eq:elbo}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta} \| \pi_{\lambda})
      + \frac{N}{M} \sum_{i=1}^M
        \log p_{\phi}(z_i \mid g(\varepsilon_i; \theta))
    \,,
\end{equation}
for a random subsample $(z_i)_{i=1}^M$ from $D$ and iid $
  (\varepsilon_i)_{i=1}^M \sim p_\varepsilon
$.
%
\todo{DSVI}

@@

In particular, natural identification of $\cplx$ and $\real^2$ implies that linear layers $
  L \colon \cplx^m \to \cplx^n
$ act upon their inputs as follows:
\begin{equation}  \label{eq:cplx-lin-op}
  {(\real^n)}^{2}
    \to {(\real^m)}^{2}
    \colon (u, v)
      \mapsto \Bigl(
        P u - Q v,
        P v + Q u
      \Bigr)
    \,,
\end{equation}
where $
  P, Q \colon \real^{n} \to \real^{m}
$ are unique operators such that $L = P + j Q$. A $\cplx$-convolutional layer with kernel
$W$ can be implemented as two $\real$-convolutions with kernels $\Re{W}$ and $\Im{W}$ using
\eqref{eq:cplx-lin-op}. Algebraic operations, trigonometric functions, are implemented
through operations in $\real^2$. Activations, considered in \citep{trabelsi_deep_2017} include
independent componentwise application of $\real$-valued non-linearities, $
  z = u + j v
    \mapsto \sigma(u) + j \sigma(v)
$, $\cplx$-differentiable exponential and trigonometric activations, and \textit{modRelu}
activation, which applies soft-thresholding to the modulus and preserves the angle: $
  z = r e^{j \phi}
    \mapsto e^{j \phi} (r - b)_+
  % z \mapsto z \bigl(
  %     1 - \tfracb{\lvert z \rvert}
  %   \bigr)_+
$ with a learnable threshold parameter $b \in \real$.

@@

  \Bigl(
    \begin{smallmatrix}

@@

    \end{smallmatrix}
  \Bigr)

@@

        \end{smallmatrix}
      \Bigr),
      \tfrac12 \Bigl(
        \begin{smallmatrix}
          \Re{(\Gamma + C)} & \Im{(C - \Gamma)} \\

@@

The entropy of a $\cplx$-Gaussian in terms of $\Gamma$ and $C$ is

@@

A $\cplx$-Gaussian vector $z$ is \textit{proper} if $z$ and $\conj{z}$ are uncorrelated, i.e.
$C = 0$. The entropy of a proper $\cplx$-Gaussian is

@@

    \tag{\ref{eq:c-gauss-entropy-derivation}'}

@@

In $\cplx$-variational dropout we use mean-field proper $\cplx$-Gaussian approximate
posterior: for a dense linear layer parameters $W \in \cplx^{n\times m}$ follow
\begin{equation}  \label{eq:c-gauss-vi}
  % (we omit the subscript $m=1$ for brevity)
  W \sim q(W)
    % = \prod_{ij} q(W_{ij})
    = \prod_{ij} \mathcal{N}^{\cplx}(W_{ij} \mid \mu_{ij}, \sigma^2_{ij}, 0)
  \,,
\end{equation}
and we use similar factorized approximation for convolutions. We propose to MAP estimates,
of the biases i.e. assume an independent fully factorized Dirac approximation.
